### 2023-08-27 Draft.tex
Altered the sizes of some images
Fixed abstract and definition

\begin{abstract}
% +++++++++++++++++++++++++
% Insert your Abstract here (one paragraph summary)
% +++++++++++++++++++++++++
\end{abstract}

@@

The two major categories of filtering in image processing are \textbf{\emph{Linear Filters}} and \textbf{\emph{Non-linear Filters}}. Linear Filtering is a simpler, yet powerful strategy for image processing, as they
only require a linear neighborhood operation to calculate the output image. For difficult scenarios however, it may not be possible to only rely on basic multiplications and sums. This is where Non-linear filters comes in handy, as they
feature concepts such as cellular automata. This article aims to discuss different types of filters, their applications and how each of them provides different benefits\footnotemark\footnotetext{Written by Jose}
\section{Linear Filters}
With \textbf{Linear Filtering}, or specifically linear spatial filtering, the function which is used to pass the image through
must be linear and shift invariant. The function $\pmb{L}$ is considered linear if there exists two constants $a$ and $b$ for any
two input pixels $f_1(m, n)$ and $f_2(m, n)$ such that

@@

also called the \emph{superposition property of linear systems}\cite{BovikActon09}. Whereas a function is considered \emph{shift invariant} when
\begin{align*}
    \implies & g(m, n) = \boldsymbol{L} [f(m, n)]\\
    \implies & g(m - p, n - q) =  \boldsymbol{L} [f(m - p, n - q)]
\end{align*} $p$ and $q$ implies a spatial shift to both the input and output pixels\cite{BovikActon09}. An intuitive way to think
of these two properties is that the function has to behave the same throughout the entire picture.

@@

\begin{minipage}{\textwidth}\begin{figure}[H]
    \centering
    \subfigure[zero]{\includegraphics[width=0.19\textwidth]{img/border_constant}}
    \subfigure[wrap]{\includegraphics[width=0.19\textwidth]{img/border_wrap}}
    \subfigure[clamp]{\includegraphics[width=0.19\textwidth]{img/border_clamp}}
    \subfigure[mirror]{\includegraphics[width=0.19\textwidth]{img/border_reflect}}
    \subfigure[extend]{\includegraphics[width=0.19\textwidth]{img/border_extend.jpg}}
    \caption{Different types of border padding}
    \label{fig:padding_borders}
\end{figure}\end{minipage}
\begin{itemize}
    \item \textbf{Zero:} Set all pixels outside the region to zero.
    \item \textbf{Constant:} This is similar to Zero, but instead a specific color is chosen to replace the out of bound pixels.
    \item \textbf{Clamp:} The edge pixels are repeated indefinitely.
    \item \textbf{(Cyclic) Wrap (Repeat or Tile):} The entire image is repeated similar to tiling it infinitely.
    \item \textbf{Mirror:} The image is mirrored accordingly on each of the edges.
    \item \textbf{Extend:} This is a combination of clamping and mirroring, where the clamped image is subtracted with the mirrored image.
\end{itemize}\label{chap:padding}

This of course not only affects the unseen pixels, but the effects can bleed into the pixels inside the frame.
\begin{minipage}{\textwidth}\begin{figure}[H]
    \centering
    \subfigure[zero]{\includegraphics[width=0.19\textwidth]{img/boxBlur_constant}}
    \subfigure[wrap]{\includegraphics[width=0.19\textwidth]{img/boxBlur_wrap}}
    \subfigure[clamp]{\includegraphics[width=0.19\textwidth]{img/boxBlur_clamp}}
    \subfigure[mirror]{\includegraphics[width=0.19\textwidth]{img/boxBlur_reflect}}
    \subfigure[extend]{\includegraphics[width=0.19\textwidth]{img/boxBlur_extend.jpg}}
    \caption{Effects of different paddings with box blur}
    \label{fig:boxBlur_borders}
\end{figure}\end{minipage}

As seen from Figure \ref{fig:boxBlur_borders}, with a constant padding, the color chosen for the constant value bleeds uniformly in the edges. The wrapped image
can be convincing at times, but drastic changes from one border to another may lead to undesirable bleeding edges. If we instead clamp the image, the repeated pixels are noticeably
present, depending on the kernel size, this may still be desireable. In this use case, mirror seems to be the most convincing. It cannot however be said, that mirror is objectively the better
method, as it depends on the scenarios and use cases (e.g. if a subject is placed in a white background, then the constant padding may be desirable).

@@

\begin{minipage}{\textwidth}\begin{figure}[H]
    \centering
    \subfigure[Pattern]{\includegraphics*[width=0.29\textwidth]{codes/pattern.png}}
    \subfigure[Sobel]{\includegraphics[width=0.29\textwidth]{img/directionSobel.png}}
    \subfigure[Prewitt]{\includegraphics[width=0.29\textwidth]{img/directionScharr.png}}
    \caption{Vertical edge detection using Sobel and Prewitt on a test pattern}\label{fig:pattern}
\end{figure}\end{minipage}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.5\linewidth]{medianfiltering.jpeg}
    \caption{The same image processed with a median filter and a Gaussian filter. Both filters remove salt and pepper noise (in varying degrees) and produce a result that more accurately corresponds with how the image should look without noise}
    \label{fig:enter-label} 
\end{figure}
Now we have seen an example median filtering, compared to Gaussian filtering, for a specific scenario, and what it does, it is time to dive into what a \textit{median filter actually is}. The algorithm works in this way: Given an input signal, in this case an image, it goes through every single pixel of the image, and replaces the value with the median value of the adjacent pixels. One can also let it go through the image whilst checking adjacent pixels in a 3px, 10px, or generally X-pixel radius around the current one, and replace it with the median of the pixels inside this X-pixel radius. The "X-pixel radius" is called a kernel. Increasing the kernel-size of the filter might lead to a "smeared" and somewhat blurry look in many scenarios.
\begin{figure}{h}
    \centering
    \includegraphics[width=1\linewidth]{median_example.jpg}
    \caption{Processing the image with a bigger kernel radius leads to less "crispness"}
    \label{fig:enter-label}
\end{figure}
\begin{figure}{h}
    \centering
    \includegraphics[width=1\linewidth]{medianfiltering.PNG}
    \caption{lllustration of how the mean filter algorithm works. The central pixel's value is changed from 80 to 40. 40 is the median of all the values within a 1px radius of the central pixel.}
    \label{fig:enter-label}
\end{figure}
However, even though median filters work relatively well for this specific scenario, the cure for cancer is yet to be found. There are (several) scenarios in which median filters might produce sub-optimal results.

@@

\subsubsection{Minimum- and Maximum Filtering}
Two more types of non-linear filters are minimum- and maximum filters. These two filters are often also called dilation and erosion filters, respectively. They are classified amongst the so called morphological filters. In many ways, the minimum and maximum filters are two sides of the same coin, they are very similar in all but one point. As the minimum filter moves along the image, processing as it should, it simply chooses the smallest value inside the current kernel and applies this value to the relevant pixel. Maximum filters work very similarly, but, as one might guess, they instead pick the largest value and assigns that to the relevant pixel. Given the two filters other names, dilation and erosion, it might not be too much of a leap to assume what an image processed by these filters would look like relative to the original. Minimum filters, or dilation filters, simply dilate the darker areas of images (since darker values tend to be assigned lower numbers). Effective if one wishes to thicken the borders of a dark-coloured letter on a light background in an image. Contrarily, the maximum filter, or erosion filter, will erode darker pixels from images. If a black pixel is surrounded by only white pixels, it will be gone after an erosion filter has been applied.
Lastly on minimum and maximum filters, a bit about binary images. What are binary images? They are in fact part of what led to the creation of morphological filters such as the here discussed minimum and maximum ones. As one might deduce from the name "binary images"  they operate with \textit{only} binary values. In other terms, each pixel can only have one of two values, black and white.
\begin{figure}[H]
    \centering
    \includegraphics[width=1\linewidth]{minmax1.png}
    \caption{Illustrates the usage of minimum and maximum filters on a grayscale image. Using the minimum filter expands the area of the figure, whereas applying a maximum filter reduces it's size}
\end{figure}

@@

\subsection{Further Non-Linear filter related techniques}
\subsubsection{Mean Shift}
The mean shift algorithm is often referred to as a "hill climbing algorithm". Given for instance a scatter plot, or such, the mean shift algorithm will locate the area of the plot where the data-point density is at it's highest, metaphorically speaking, a hill (densest data cluster). Roughly, the algorithm will move (shift) from data point to data point based on the regional mean (mean of current cluster). What is considered in calculating the mean may vary from implementation to implementation. Some times calculating a weighted mean may yield more accurate results, depending on what one is searching for.
How can one apply this concept in image processing? Mostly in image segmentation. As mean shift is quite prolific at finding and recognizing clusters, it is also quite good at recognizing "clusters" in images, i.e. different parts of an image. Segmenting an image, categorizing different parts of it, can allow you to perhaps apply different filtering algorithms for different areas of the image, maybe help with edge detection, and in general help preserve the intended look and data of the image. Furthermore it has applications in for instance computer vision and pattern recognition. Important (sub-)fields of image processing.
\begin{figure}[H]
    \centering
    \includegraphics[width=1\linewidth]{meanshift.jpeg}
    \caption{Four different images, all segmented by a mean shift algorithm. The segmentation enables one to analyze different areas in different ways. Maybe one would rather filter noise away from water-areas or different coloured areas with different filtering techniques}
    \label{fig:enter-label}
\end{figure}
\begin{figure}[H]
    \centering
    \includegraphics[width=1\linewidth]{scattergraph.png}
    \caption{A typical graph for which the mean shift algorithm could be applied. The clusters can be viewed as hills. Increased data point density would imply a "steeper hill". One can roughly recognize the outline of three main clusters here}
    \label{fig:enter-label}
\end{figure}

>>>>

\begin{abstract}
    Linear \textit{(spatial)} Filters are operators for images that produces a desirable output by means
    of kernels that is applied equally throughout an image. Calculating these filters by means of convolution or correlation are typically used, but
    can be optimized by separating the kernels, or with a summed area table. With different operators, such as the Sobel, Gaussian and Laplacian Kernels, these filters
    can be used for applications ranging from smoothing to edge detection. Each kernels have their own strengths in regards to quality and complexity. Non-linear filter are not bounded by the linear property of its counterpart.
    They provide filters with higher complexities, which in turns provide better results and can solve problems that linear filters and can extract information with better context, such as removing the salt and pepper noise, or only
    highlighting major edges relevant to the user. Methods such as bilateral filtering, median filtering can aid in removing noise. Further applications such as image restoration and pattern recognition can also be achieved by means of non-linear filters.
\end{abstract}

@@

The two major categories of filtering in image processing are \textbf{\emph{Linear Filters}} and \textbf{\emph{Non-linear Filters}}. Linear Filtering is a simpler, yet powerful strategy for image processing, as they
only require a linear neighborhood operation to calculate the output image. For difficult scenarios however, it may not be possible to only rely on basic multiplications and sums. This is where Non-linear filters comes in handy, as they
feature concepts such as cellular automata. This article aims to discuss different types of filters, their applications and how each of them provides different benefits\footnotemark\footnotetext{Written by Jose}
\section{Linear Filters}
With \textbf{Linear Filtering}, or specifically linear spatial filtering, the function which is used to pass the image through
must be linear and shift invariant. The function $\pmb{L}$ is considered linear if for any two constants $a$ and $b$, and for any
two input pixels $f_1(m, n)$ and $f_2(m, n)$ such that

@@

This is also called the \emph{superposition property of linear systems}\cite{BovikActon09}. Whereas a function is considered \emph{shift invariant} when
\begin{align*}
    \implies & g(m, n) = \boldsymbol{L} [f(m, n)]\\
    \implies & g(m - p, n - q) =  \boldsymbol{L} [f(m - p, n - q)]
\end{align*} $p$ and $q$ implies a spatial shift to both the input and output pixels\cite{BovikActon09}. An intuitive way to think
of these two properties is that the function has to behave the same throughout the entire picture.

@@

\begin{minipage}{\textwidth}\begin{figure}[H]
    \captionsetup{justification=centering}
    \centering
    \subfigure[zero]{\includegraphics[width=0.19\textwidth]{img/border_constant}}
    \subfigure[wrap]{\includegraphics[width=0.19\textwidth]{img/border_wrap}}
    \subfigure[clamp]{\includegraphics[width=0.19\textwidth]{img/border_clamp}}
    \subfigure[mirror]{\includegraphics[width=0.19\textwidth]{img/border_reflect}}
    \subfigure[extend]{\includegraphics[width=0.19\textwidth]{img/border_extend.jpg}}
    \caption{Different types of border padding}
    \label{fig:padding_borders}
\end{figure}\end{minipage}
\begin{itemize}
    \item \textbf{Zero:} Set all pixels outside the region to zero.
    \item \textbf{Constant:} This is similar to Zero, but instead a specific color is chosen to replace the out of bound pixels.
    \item \textbf{Clamp:} The edge pixels are repeated indefinitely.
    \item \textbf{(Cyclic) Wrap (Repeat or Tile):} The entire image is repeated similar to tiling it infinitely.
    \item \textbf{Mirror:} The image is mirrored accordingly on each of the edges.
    \item \textbf{Extend:} This is a combination of clamping and mirroring, where the clamped image is subtracted with the mirrored image.
\end{itemize}\label{chap:padding}

This of course not only affects the unseen pixels, but the effects can bleed into the pixels inside the frame.
\begin{minipage}{\textwidth}\begin{figure}[H]
    \captionsetup{justification=centering}
    \centering
    \subfigure[zero]{\includegraphics[width=0.19\textwidth]{img/boxBlur_constant}}
    \subfigure[wrap]{\includegraphics[width=0.19\textwidth]{img/boxBlur_wrap}}
    \subfigure[clamp]{\includegraphics[width=0.19\textwidth]{img/boxBlur_clamp}}
    \subfigure[mirror]{\includegraphics[width=0.19\textwidth]{img/boxBlur_reflect}}
    \subfigure[extend]{\includegraphics[width=0.19\textwidth]{img/boxBlur_extend.jpg}}
    \caption{Effects of different paddings with box blur}
    \label{fig:boxBlur_borders}
\end{figure}\end{minipage}

As seen from Figure \ref{fig:boxBlur_borders}, with a constant padding, the color chosen for the constant value bleeds uniformly in the edges. The wrapped image
can be convincing at times, but drastic changes from one border to another may lead to undesirable bleeding edges. If we instead clamp the image, the repeated pixels are noticeably
present, depending on the kernel size, this may still be desirable. In this use case, mirror seems to be the most convincing. It cannot however be said, that mirror is objectively the better
method, as it depends on the scenarios and use cases (e.g. if a subject is placed in a white background, then the constant padding may be desirable).

@@

\begin{minipage}{\textwidth}\begin{figure}[H]
    \captionsetup{justification=centering}
    \centering
    \subfigure[Pattern]{\includegraphics*[width=0.29\textwidth]{codes/pattern.png}}
    \subfigure[Sobel]{\includegraphics[width=0.29\textwidth]{img/directionSobel.png}}
    \subfigure[Prewitt]{\includegraphics[width=0.29\textwidth]{img/directionPrewitt.png}}
    \caption{Vertical edge detection using Sobel and Prewitt on a test pattern}\label{fig:pattern}
\end{figure}\end{minipage}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.5\linewidth]{medianfiltering.jpeg}
    \caption{The same image processed with a median filter and a Gaussian filter. Both filters remove salt and pepper noise (in varying degrees) and produce a result that more accurately corresponds with how the image should look without noise}
    \label{fig:enter-label} 
\end{figure}
Now we have seen an example median filtering, compared to Gaussian filtering, for a specific scenario, and what it does, it is time to dive into what a \textit{median filter actually is}. The algorithm works in this way: Given an input signal, in this case an image, it goes through every single pixel of the image, and replaces the value with the median value of the adjacent pixels. One can also let it go through the image whilst checking adjacent pixels in a 3px, 10px, or generally X-pixel radius around the current one, and replace it with the median of the pixels inside this X-pixel radius. The "X-pixel radius" is called a kernel. Increasing the kernel-size of the filter might lead to a "smeared" and somewhat blurry look in many scenarios.
\begin{figure}{h}
    \centering
    \includegraphics[width=0.75\linewidth]{median_example.jpg}
    \caption{Processing the image with a bigger kernel radius leads to less "crispness"}
    \label{fig:enter-label}
\end{figure}
\begin{figure}{h}
    \centering
    \includegraphics[width=1\linewidth]{medianfiltering.PNG}
    \caption{lllustration of how the mean filter algorithm works. The central pixel's value is changed from 80 to 40. 40 is the median of all the values within a 1px radius of the central pixel.}
    \label{fig:enter-label}
\end{figure}
However, even though median filters work relatively well for this specific scenario, the cure for cancer is yet to be found. There are (several) scenarios in which median filters might produce sub-optimal results.

@@

\subsubsection{Minimum- and Maximum Filtering}
Two more types of non-linear filters are minimum- and maximum filters. These two filters are often also called dilation and erosion filters, respectively. They are classified amongst the so called morphological filters. In many ways, the minimum and maximum filters are two sides of the same coin, they are very similar in all but one point. As the minimum filter moves along the image, processing as it should, it simply chooses the smallest value inside the current kernel and applies this value to the relevant pixel. Maximum filters work very similarly, but, as one might guess, they instead pick the largest value and assigns that to the relevant pixel. Given the two filters other names, dilation and erosion, it might not be too much of a leap to assume what an image processed by these filters would look like relative to the original. Minimum filters, or dilation filters, simply dilate the darker areas of images (since darker values tend to be assigned lower numbers). Effective if one wishes to thicken the borders of a dark-coloured letter on a light background in an image. Contrarily, the maximum filter, or erosion filter, will erode darker pixels from images. If a black pixel is surrounded by only white pixels, it will be gone after an erosion filter has been applied.
Lastly on minimum and maximum filters, a bit about binary images. What are binary images? They are in fact part of what led to the creation of morphological filters such as the here discussed minimum and maximum ones. As one might deduce from the name "binary images"  they operate with \textit{only} binary values. In other terms, each pixel can only have one of two values, black and white.
\begin{figure}[H]
    \centering
    \includegraphics[width=0.75\linewidth]{minmax1.png}
    \caption{Illustrates the usage of minimum and maximum filters on a grayscale image. Using the minimum filter expands the area of the figure, whereas applying a maximum filter reduces it's size}
\end{figure}

@@

\subsection{Further Non-Linear filter related techniques}
\subsubsection{Mean Shift}
The mean shift algorithm is often referred to as a "hill climbing algorithm". Given for instance a scatter plot, or such, the mean shift algorithm will locate the area of the plot where the data-point density is at it's highest, metaphorically speaking, a hill (densest data cluster). Roughly, the algorithm will move (shift) from data point to data point based on the regional mean (mean of current cluster). What is considered in calculating the mean may vary from implementation to implementation. Some times calculating a weighted mean may yield more accurate results, depending on what one is searching for.
How can one apply this concept in image processing? Mostly in image segmentation. As mean shift is quite prolific at finding and recognizing clusters, it is also quite good at recognizing "clusters" in images, i.e. different parts of an image. Segmenting an image, categorizing different parts of it, can allow you to perhaps apply different filtering algorithms for different areas of the image, maybe help with edge detection, and in general help preserve the intended look and data of the image. Furthermore it has applications in for instance computer vision and pattern recognition. Important (sub-)fields of image processing.
\begin{figure}[H]
    \centering
    \includegraphics[width=1\linewidth]{meanshift.jpeg}
    \caption{Four different images, all segmented by a mean shift algorithm. The segmentation enables one to analyze different areas in different ways. Maybe one would rather filter noise away from water-areas or different coloured areas with different filtering techniques}
    \label{fig:enter-label}
\end{figure}
\begin{figure}[H]
    \centering
    \includegraphics[width=0.75\linewidth]{scattergraph.png}
    \caption{A typical graph for which the mean shift algorithm could be applied. The clusters can be viewed as hills. Increased data point density would imply a "steeper hill". One can roughly recognize the outline of three main clusters here}
    \label{fig:enter-label}
\end{figure}
